---
title: "Homework 3"
author: "Kevin Patterson"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continous.colour = "viridis",
  ggplot2.continous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete  = scale_fill_viridis_d
```

# Problem 1
```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user and order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes.

How many aisles, and which are most items from?
```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
There are 134 aisles and the most items ordered are from fresh vegetables aisle (n=150,609), fresh fruits aisle (n=150,473), followed by packaged vegetables and fruits aisle (n=78,493).

Let's make a plot
```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
Let's make a table
```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```
Let's make a table on Apples vs. Ice Cream.
```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

# Problem 2
Load, tidy, wrangle data.
```{r}
accel_df = read_csv("./accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    cols = activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_counts") %>%
  mutate(week_type=case_when(day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "Weekday",
                           day %in% c("Saturday","Sunday") ~ "Weekend")
         )
```
Tidied dataset
```{r}
accel_df%>%tibble()
```
This new dataset contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. There are 50,400 observations of activity counts that correspond to the minute they were recorded. There are 6 variables that include 5 weeks of accelerometer observations, a `day_id` variable that corresponds to each day of the study (n = 35 days), and a `week_type` variable that indicates whether an observation was on a weekday or on the weekend.

Aggregate dataset across minutes to create a total activity variable for each day.
```{r}
accel_aggdf = accel_df %>%
  group_by(week, day_id, day) %>%
  summarize(activity_sum = sum(activity_counts)) %>%
  arrange(week) %>%
  rename("study day" = day_id,
         "total activity" = activity_sum)
  knitr::kable(accel_aggdf)
```

The aggregated accelerometer dataset doesn't seem to follow any particular trend, I arranged by `week` initially to see if there was increased or decreased activity as each week progressed, but there is no such pattern observed. Then, I saw that higher `total activity` values around Friday to Sunday, but after week 3 this pattern did not hold up. There are very low values for `study day` 24 and 31, which might indicate that less observations were taken on these two days resulting in their outlier behavior. I also arranged by descending order of `total activity` (not shown), but after reviewing the corresponding variables (i.e., `week`, `day`, `study day`) no such pattern was observed. 

Let's make a single panel plot that shows 24-hour activity time courses for each day (using viridis from library setup)
```{r}
accel_actdf = accel_df %>%
  mutate(minute = as.numeric(minute),
         hour = minute / 60) %>%
  group_by(day, hour) %>%
  summarize(activity_sum = sum(activity_counts)) %>%
  ggplot(aes(x = hour, y = activity_sum)) +
  geom_point(aes(color = day)) + geom_line(aes(color = day)) +
  viridis::scale_color_viridis(name = "day",
                               discrete = "TRUE",
                               option = "plasma") +
  ggtitle("24-Hour Activity Time Courses for each day") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Hour",
       y = "Activity Counts") +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24), 
                     labels = c("midnight", "1AM", "2AM", "3AM", "4AM", "5AM", "6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "noon", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM", "11PM", "midnight")) +
  scale_y_continuous(
    breaks = c(0, 50000, 100000, 150000, 200000, 250000, 300000, 350000),
    labels = c("0", "50,000", "100,0000", "150,000", "200,000", "250,000", "300,000", "350,000"))
accel_actdf
```
The activity counts for the 63 year-old male are highest in the evenings around 8PM to 9PM and then followed by counts at 11AM across all days. As mentioned in the last question, it seemed that highest counts of activity were around the weekend and the plot better illustrates the highest activity counts on Fridays, Sundays, and Mondays. The male does activity more around the mornings and evenings.

# Problem 3

Load NY NOAA dataset
```{r}
library(p8105.datasets)
data("ny_noaa")
```

```{r}
ny_noaa%>%tibble()
```
This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. This dataset has 2,595,176 observatons and 7 variables that include a weather station id, date of observation, (tenths of mm), snowfall (mm), snow depth (mm), maximum and minimum temperatures (tenths of degree C). There are missing data for `prcp`, `snow`, `snwd`, `tmax`, and `tmin` variables. The dataset has recorded data from January 1, 1981 through Decmber 31, 2010 from the National Oceanic and Atmospheric Association of the National Centers for Environmental Information.
